### 2020-11-10 manuscript/manuscript.rmd
change Choanoflagellata to Choanoflagellatea

Sampling of outgroup taxa (non-animals, in this case) is critical to phylogenetic rooting questions, since the node where the outgroup subtree attaches to the ingroup subtree is the root of the ingroup. There has therefore been extensive focus on improving outgroup sampling when testing phylogenetic hypotheses about rooting [@graham2002rooting]. Most studies addressing the animal root have removed more distantly related outgroup taxa in some analyses to explore the effect of outgroup selection to ingroup topology [@ryan2013genome; @pisani2015genomic]. Three progressively more inclusive clades have often been investigated: Choanozoa (animals plus most closely related Choanoflagellata), Holozoa (Choanozoa plus more distantly related Holozoa), and Opisthokonta (Holozoa + Fungi).

@@

All IQ-TREE analyses, apart from unresolved analyses (for Moroz2014_3d and all Nosenko2013 matrices), supported Ctenophora-sister (Fig. 3B). No IQ-TREE analyses supported Porifera-sister, including those that restrict outgroup sampling to Choanoflagellata and use models with site-heterogeneous equilibrium frequencies (Fig. 3B lower right; Supplementary Table 2), the conditions under which published PhyloBayes CAT analyses recover strong support for Porifera-sister (Fig. 3A). To further verify this difference in a controlled manner we reran PhyloBayes analyses with CAT, using both Poisson and GTR substitution matrices, for some matrices that had led to support for Porifera-sister in published analyses. Consistent with published results, some of these supported Porifera-sister.

@@

#### Outgroup taxa sampling with C60 and CAT models
Because different choices of outgroups could also affect phylogenetic inference as suggested in previous analyses, we parsed the full data matrices into three different types of outgroups: Choanozoa, Holozoa and Opisthokonta. These datasets include the same set of genes but differ in the composition of outgroup species. Choanozoa only includes choanofagellates as outgroup; Holozoa also includes more distantly related Holozoans; Opistokonta also includes Fungi. For each Choanozoa data matrice, both C60 models in IQ-TREE and Poisson+CAT models in PhyloBayes were conducted. The maximum likelihood analysis was performed using the best-fit substitution model identified as above and nodal support was assessed with 1000 ultrafast bootstrap replicates using IQ-TREE. Moreover, Bayesian inference with the site-heterogeneous Poisson+CAT model was done with PhyloBayes-MPI v1.8. To minimize computational burden, GTR+CAT models were only performed in the representative Choanozoa matrices from Philippe2009, Ryan2013_est and Whelan2017_full.

@@

Pick *et al.* [@Pick:2010eb] sought to test whether Ctenophora-sister was an artefact of insufficient taxon sampling. They added new and additional published sequence data to the 64-taxon matrix of Dunn *et al.* [@Dunn:2008ky]. The new taxa included 12 sponges, 1 ctenophore, 5 cnidarians, and *Trichoplax*. They further modified the matrix by removing 2,150 sites that were poorly sampled or aligned. They considered two different sets of outgroups: Choanoflagellata (resulting in Choanozoa) and the same sampling as Dunn *et al.* (resulting in Opisthokonta).

>>>>

Sampling of outgroup taxa (non-animals, in this case) is critical to phylogenetic rooting questions, since the node where the outgroup subtree attaches to the ingroup subtree is the root of the ingroup. There has therefore been extensive focus on improving outgroup sampling when testing phylogenetic hypotheses about rooting [@graham2002rooting]. Most studies addressing the animal root have removed more distantly related outgroup taxa in some analyses to explore the effect of outgroup selection to ingroup topology [@ryan2013genome; @pisani2015genomic]. Three progressively more inclusive clades have often been investigated: Choanozoa (animals plus most closely related Choanoflagellatea), Holozoa (Choanozoa plus more distantly related Holozoa), and Opisthokonta (Holozoa + Fungi).

@@

All IQ-TREE analyses, apart from unresolved analyses (for Moroz2014_3d and all Nosenko2013 matrices), supported Ctenophora-sister (Fig. 3B). No IQ-TREE analyses supported Porifera-sister, including those that restrict outgroup sampling to Choanoflagellatea and use models with site-heterogeneous equilibrium frequencies (Fig. 3B lower right; Supplementary Table 2), the conditions under which published PhyloBayes CAT analyses recover strong support for Porifera-sister (Fig. 3A). To further verify this difference in a controlled manner we reran PhyloBayes analyses with CAT, using both Poisson and GTR substitution matrices, for some matrices that had led to support for Porifera-sister in published analyses. Consistent with published results, some of these supported Porifera-sister.

@@

#### Outgroup taxa sampling with C60 and CAT models
Because different choices of outgroups could also affect phylogenetic inference as suggested in previous analyses, we parsed the full data matrices into three different types of outgroups: Choanozoa, Holozoa and Opisthokonta. These datasets include the same set of genes but differ in the composition of outgroup species. Choanozoa only includes Choanoflagellatea outgroup; Holozoa also includes more distantly related Holozoans; Opistokonta also includes Fungi. For each Choanozoa data matrice, both C60 models in IQ-TREE and Poisson+CAT models in PhyloBayes were conducted. The maximum likelihood analysis was performed using the best-fit substitution model identified as above and nodal support was assessed with 1000 ultrafast bootstrap replicates using IQ-TREE. Moreover, Bayesian inference with the site-heterogeneous Poisson+CAT model was done with PhyloBayes-MPI v1.8. To minimize computational burden, GTR+CAT models were only performed in the representative Choanozoa matrices from Philippe2009, Ryan2013_est and Whelan2017_full.

@@

Pick *et al.* [@Pick:2010eb] sought to test whether Ctenophora-sister was an artefact of insufficient taxon sampling. They added new and additional published sequence data to the 64-taxon matrix of Dunn *et al.* [@Dunn:2008ky]. The new taxa included 12 sponges, 1 ctenophore, 5 cnidarians, and *Trichoplax*. They further modified the matrix by removing 2,150 sites that were poorly sampled or aligned. They considered two different sets of outgroups: Choanoflagellatea (resulting in Choanozoa) and the same sampling as Dunn *et al.* (resulting in Opisthokonta).
